`State and Parameter Estimation for Spatio-Temporal Neural Fields' or `A Data Driven Framework for Patient-Specific Neural Field Modelling'

### 2010-06-20 ltx/Journal of Neuroengineering/StateAndParameterEstimationNeuralFields_JoNE.tex
changes to appendix

\section{Derivation for equations~\ref{eq:GaussianFT} and \ref{eq:WidthFrequencyRelationship} }\label{ap:FrequencyAnalysis}
Applying the \textit{n}-dimensional Fourier transform \cite{Arsac1966} to a \textit{n}-dimensional Gaussian centered at the origin (see equation~\ref{eq:BasisFunctionAtOrigin}) yields

@@

&\times& \mathrm{exp}\left(-\sigma_{\phi}^2\pi^2\boldsymbol\nu^\top \boldsymbol\nu\right)

@@

=\left(\pi\sigma_{\phi}^2\right)^{\frac{n}{2}}&&

@@

substituting \ref{eq:IntegralOfGaussian} in \ref{eq:AppendixGaussianFT} gives another scaled  Gaussian 

@@

   \boldsymbol\Phi(\boldsymbol\nu)=(\pi\sigma_{\phi}^2)^{\frac{n}{2}}\mathrm{exp}(-\sigma_{\phi}^2\pi^2\boldsymbol\nu^\top \boldsymbol\nu)

@@

substituting $\sigma_{\nu}^2=\frac{1}{\pi^2\sigma_{\phi}^2}$ gives 

@@

In this appendix we provide derivations for the convolution and the inner product of Gaussian basis functions. These derivations are used in the analytic calculation of $\boldsymbol\Psi(\mathbf{r})$. Consider two Gaussian basis functions

@@

 \phi_i(\mathbf r)=\mathrm{exp}\left({-\frac{1}{\sigma_i^2} (\mathbf r-\boldsymbol \mu_i)^\top(\mathbf r-\boldsymbol \mu_i})\right)

@@

@@

\phi_j(\mathbf r)=\mathrm{exp}\left({-\frac{1}{\sigma_j^2} (\mathbf r-\boldsymbol \mu_j)^\top(\mathbf r-\boldsymbol \mu_j})\right)

@@

First we derive the convolution of two Gaussian basis functions assuming $\boldsymbol\mu_i=\boldsymbol\mu_j=\mathbf 0$. Convolution theorem states that the Fourier transform of a convolution is the product of Fourier transforms. Using equation \ref{eq:FTofGaussian} we have

@@

\boldsymbol\Phi_i(\boldsymbol\nu)\boldsymbol\Phi_j(\boldsymbol\nu)&=&\left(\pi\sigma_i^2\sigma_j^2\right)^{\frac{n}{2}}\left(\frac{1}{\sigma_i^2+\sigma_j^2}\right)^{\frac{n}{2}} \nonumber \\
&\times& (\pi(\sigma_i^2+\sigma_j^2))^{\frac{n}{2}}\mathrm{exp}\left(-(\sigma_i^2+\sigma_j^2)\pi^2\boldsymbol\nu^\top\boldsymbol\nu\right),

@@

taking the inverse Fourier transform we can write the convolution as

@@

 (\phi_i*\phi_j)(\mathbf r)=\left(\frac{\pi\sigma_i^2\sigma_j^2}{\sigma_i^2+\sigma_j^2}\right)^{\frac{n}{2}}\mathrm{exp}\left({-\frac{1}{\sigma_i^2+\sigma_j^2} \mathbf r^\top\mathbf r}\right).

@@

In the case where $\boldsymbol\mu_i\neq\boldsymbol\mu_j\neq\mathbf 0$ equation \ref{eq:ConvoftwoGaussian} becomes

@@

 (\phi_i*\phi_j)(\mathbf r)=\left(\frac{\pi\sigma_i^2\sigma_j^2}{\sigma_i^2+\sigma_j^2}\right)^{\frac{n}{2}}\mathrm{exp}\left({-\frac{1}{\sigma_i^2+\sigma_j^2} (\mathbf r-\boldsymbol\mu)^\top(\mathbf r-\boldsymbol\mu)}\right).

@@

where $\boldsymbol\mu=\boldsymbol\mu_i+\boldsymbol\mu_j$.

The inner product of two Gaussian basis functions is defined as
\begin{equation}\label{eq:InofGaussianDefinition}
\phi_i\otimes\phi_j=\int_{\mathcal R^n}\phi_i(\mathbf r)\phi_j(\mathbf r)d\mathbf r
\end{equation}
A Gaussian basis function is symmetric with respect to its center, therefore
\begin{equation}\label{eq:GaussianSymmetry}
 \phi_i(\boldsymbol \mu_i+\mathbf r)= \phi_i(\boldsymbol \mu_i-\mathbf r)
\end{equation}
A direct consequence of equation \ref{eq:GaussianSymmetry} is
\begin{equation}\label{eq:ConsequenceofGaussianSymmetry}
 \phi_i(\mathbf r)= \phi_i(2\boldsymbol \mu_i-\mathbf r)
\end{equation}
substituting equation \ref{eq:ConsequenceofGaussianSymmetry} in equation \ref{eq:InofGaussianDefinition} we have 
\begin{eqnarray}
\phi_i\otimes\phi_j&=&\int_{\mathcal R^n}\phi_i(\mathbf r)\phi_j(\mathbf r)d\mathbf r\\
&=&\int_{\mathcal R^n}\phi_i(2\boldsymbol \mu_i-\mathbf r)\phi_j(\mathbf r)d\mathbf r\\
&=&(\phi_i*\phi_j)(2\boldsymbol \mu_i)
\end{eqnarray}
evaluating equation \ref{eq:ConvoftwoGaussianNonzeroMean} at $\mathbf r=2\boldsymbol\mu_i $ gives
\begin{equation}
 \phi_i\otimes\phi_j=\left(\frac{\pi\sigma_i^2\sigma_j^2}{\sigma_i^2+\sigma_j^2}\right)^{\frac{n}{2}}\mathrm{exp}\left({-\frac{1}{\sigma_i^2+\sigma_j^2} (\boldsymbol\mu_i-\boldsymbol\mu_j)^\top(\boldsymbol\mu_i-\boldsymbol\mu_j)}\right).
\end{equation}
\section{Notation}

>>>>

\section{Calculation of Spatial Cut-off Frequency of Field Basis Functions}\label{ap:FrequencyAnalysis}
This Appendix provides the derivation of the field basis function cut-off frequency as a function of the basis function width (equations~\ref{eq:GaussianFT} and \ref{eq:WidthFrequencyRelationship} in the main text). Applying the \textit{n}-dimensional Fourier transform \cite{Arsac1966} to a \textit{n}-dimensional Gaussian centred at the origin (see equation~\ref{eq:BasisFunctionAtOrigin}) yields

@@

&\times& \mathrm{exp}\left(-\sigma_{\phi}^2\pi^2\boldsymbol\nu^\top \boldsymbol\nu\right),

@@

=\left(\pi\sigma_{\phi}^2\right)^{\frac{n}{2}}.&&

@@

Substituting equation~\ref{eq:IntegralOfGaussian} into equation~\ref{eq:AppendixGaussianFT} gives the scaled Gaussian 

@@

   \boldsymbol\Phi(\boldsymbol\nu)=(\pi\sigma_{\phi}^2)^{\frac{n}{2}}\mathrm{exp}(-\sigma_{\phi}^2\pi^2\boldsymbol\nu^\top \boldsymbol\nu).

@@

Substituting $\sigma_{\nu}^2=\frac{1}{\pi^2\sigma_{\phi}^2}$ gives 

@@

In this appendix we provide derivations for the convolution and the inner product of Gaussian basis functions. These derivations are used in the analytic calculation of $\boldsymbol\Psi(\mathbf{r})$. Consider the field Gaussian basis function

@@

 \phi_i(\mathbf r)=\mathrm{exp}\left({-\frac{1}{\sigma_{\phi}^2} (\mathbf r-\boldsymbol \mu_i)^\top(\mathbf r-\boldsymbol \mu_i})\right)

@@

and the connectivity kernel basis function

@@

\psi_j(\mathbf r)=\mathrm{exp}\left({-\frac{1}{\sigma_{\psi_j}^2} \mathbf{r}^\top\mathbf{r}}\right)

@@

We will proceed in the derivation of the convolution assuming the field basis function is positioned at the origin ($\boldsymbol\mu_i=\mathbf 0$). The convolution theorem states that the Fourier transform of a convolution is the product of Fourier transforms. Using the derivation of the Fourier transform of a \emph{n}-dimensional Gaussian from equation~\ref{eq:FTofGaussian} of \ref{ap:FrequencyAnalysis} we have

@@

\mathcal{F}\left(\phi_i(\mathbf{r})*\psi_j(\mathbf{r}'-\mathbf{r})\right) &=& \left(\pi\sigma_{\phi}^2\sigma_{\psi_j}^2\right)^{\frac{n}{2}}\left(\frac{1}{\sigma_{\phi}^2+\sigma_{\psi_j}^2}\right)^{\frac{n}{2}} \nonumber \\
&\times& (\pi(\sigma_{\phi}^2+\sigma_{\psi_j}^2))^{\frac{n}{2}}\mathrm{exp}\left(-(\sigma_{\phi}^2+\sigma_{\psi_j}^2)\pi^2\boldsymbol\nu^\top\boldsymbol\nu\right),

@@

where $\mathcal{F}$ denotes the Fourier transform. Taking the inverse Fourier transform, the convolution is

@@

 \phi_i(\mathbf{r})*\psi_j(\mathbf{r}'-\mathbf{r}) = \left(\frac{\pi\sigma_{\phi}^2\sigma_{\psi_j}^2}{\sigma_{\phi}^2+\sigma_{\psi_j}^2}\right)^{\frac{n}{2}}\mathrm{exp}\left({-\frac{1}{\sigma_{\phi}^2+\sigma_{\psi_j}^2} \mathbf r^\top\mathbf r}\right).

@@

In the case where $\boldsymbol\mu_i\neq\mathbf 0$ equation \ref{eq:ConvoftwoGaussian} becomes

@@

 (\phi_i*\phi_j)(\mathbf r)=\left(\frac{\pi\sigma_{\phi}^2\sigma_{\psi_j}^2}{\sigma_{\phi}^2+\sigma_{\psi_j}^2}\right)^{\frac{n}{2}}\mathrm{exp}\left({-\frac{1}{\sigma_{\phi}^2+\sigma_{\psi_j}^2} (\mathbf r-\boldsymbol\mu_i)^\top(\mathbf r-\boldsymbol\mu_i)}\right).

@@

The inner product of two Gaussian basis functions is defined as
\begin{equation}\label{eq:InofGaussianDefinition}
\phi_i\otimes\phi_j=\int_{\mathcal R^n}\phi_i(\mathbf r)\phi_j(\mathbf r)d\mathbf r
\end{equation}
A Gaussian basis function is symmetric with respect to its centre, therefore
\begin{equation}\label{eq:GaussianSymmetry}
 \phi_i(\boldsymbol \mu_i+\mathbf r)= \phi_i(\boldsymbol \mu_i-\mathbf r)
\end{equation}
A direct consequence of equation \ref{eq:GaussianSymmetry} is
\begin{equation}\label{eq:ConsequenceofGaussianSymmetry}
 \phi_i(\mathbf r)= \phi_i(2\boldsymbol \mu_i-\mathbf r)
\end{equation}
substituting equation \ref{eq:ConsequenceofGaussianSymmetry} in equation \ref{eq:InofGaussianDefinition} we have 
\begin{eqnarray}
\phi_i\otimes\phi_j&=&\int_{\mathcal R^n}\phi_i(\mathbf r)\phi_j(\mathbf r)d\mathbf r\\
&=&\int_{\mathcal R^n}\phi_i(2\boldsymbol \mu_i-\mathbf r)\phi_j(\mathbf r)d\mathbf r\\
&=&(\phi_i*\phi_j)(2\boldsymbol \mu_i)
\end{eqnarray}
evaluating equation \ref{eq:ConvoftwoGaussianNonzeroMean} at $\mathbf r=2\boldsymbol\mu_i $ gives
\begin{equation}
 \phi_i\otimes\phi_j=\left(\frac{\pi\sigma_i^2\sigma_j^2}{\sigma_i^2+\sigma_j^2}\right)^{\frac{n}{2}}\mathrm{exp}\left({-\frac{1}{\sigma_i^2+\sigma_j^2} (\boldsymbol\mu_i-\boldsymbol\mu_j)^\top(\boldsymbol\mu_i-\boldsymbol\mu_j)}\right).
\end{equation}
\section{Notation}
